# Exploratory data analysis 

## Univariate analysis

In this first part of our EDA, we will explore each variable on its own, before looking at the relationship between different variables.

### Plots of each variable over time: 

```{r}
#Flights plot 
ts_data %>% autoplot(Flights) +
  labs(title = "Flights over time", x = "Date", y = "Number of Flights")

#Price plot
ts_data %>% autoplot(Price) +
  labs(title = "Brent Oil Prices 2019 onwards", x = "Date", y = "Price")

#temp plot
ts_data %>% autoplot(temp) +
  labs(title = "Temperature over time", x = "Date", y = "Temperature in Celsius")

#precip plot
ts_data %>% autoplot(precip) +
  labs(title = "Precipitation over time", x = "Date", y = "Precipitation")

#snow plot
ts_data %>% autoplot(snow) +
  labs(title = "Snow over time", x = "Date", y = "Snow")

#windspeed plot
ts_data %>% autoplot(windspeed) +
  labs(title = "Windspeed over time", x = "Date", y = "Windspeed")

#visibility plot
ts_data %>% autoplot(visibility) +
  labs(title = "Visibility over time", x = "Date", y = "Visibility")

#humidity plot
ts_data %>% autoplot(humidity) +
  labs(title = "Humidity over time", x = "Date", y = "Humidity")


```
Here are some observations based on the data-over-time graphs above: 

- Number of Flights: There is a clear trend, and some possible seasonality. Also, the covid flight-restriction drop is extremely clear. After that, the number of flights seems to be steadily increasing (with some cycles) towards to its pre-covid level. However, from this graph we can argue that there tend to be more flights in the summer compared to the winter. A big factor causing this is obviously school holidays in the summer.
- Brent oil price: There is a clear trend, but not an obvious seasonality in the data at first glance. Further analysis is needed to determine the specifics.
- Temperature: Extremely seasonal data, without a clear trend. Always peaking in the summer months for the highest temperatures and in the winter months for the lowest ones.
- Precipitation: Difficult to interpret at this point, but some seasonality can be seen, with less precipitation in the summer months, and more in the winter months.
- Snow: Very bizarre that there's no snow at all in 2020, and then a huge peak in 2021. I need to analyze in more depth to see what's going on, but since there are so many zeros for most of the year, I'm not sure if snow as a variable will help forecast anything. 
- Windspeed: Very clear seasonality, with perhaps a downward trend for 2022-2023. 
- Visibility: Clear seasonality with the best visibility in the summer and worst in the winter. Also, the data is looking weird after 2022...
- Humidity: Clear seasonality with more humidity in the winter compared to the summer months.

### Specific plots for interesting variables 

```{r}

```



## Bivariate analysis

In this section of our EDA, we will explore the relationships between our variables, which will help us decide which to keep for our final forecasting analysis. 