## Models using Weather variables

Here's what our flights-weather dataset looks like: 

```{r}
Full_Data = Flights_Holidays %>% as_tsibble(index = date)
Full_Data_Until_Covid = Full_Data[Full_Data$date <= as.Date("2020-01-31"),] %>% as_tsibble(index= date)

datatable(Full_Data, options = list(pageLength = 6))

weather_train_set = Full_Data[Full_Data$date < as.Date("2023-03-01"),]
weather_test_set = Full_Data[Full_Data$date >= as.Date("2023-03-01"),]
weather_train_covid = Full_Data_Until_Covid[Full_Data_Until_Covid$date < as.Date("2020-01-01"),]
weather_test_covid = Full_Data_Until_Covid[Full_Data_Until_Covid$date >= as.Date("2020-01-01"),]
subset_weather = weather_train_set[weather_train_set$date >= as.Date("2022-11-01"),]
subset_weather_covid = weather_train_covid[weather_train_covid$date >= as.Date("2019-11-01"),]

```

### Forecasting Predictors with ETS 

```{r ETS_temperature_with_covid}

#Modelling ets using additive seasonal component
ets_temp <- weather_train_set %>%
  model(ETS(temp)) 

#Getting the Accuracy scores for model
report(ets_temp)

#Forecasting using the model, forcating one month
forecast_ets_temp <- ets_temp %>% forecast(h = 30)

#Graphing the forecast
ets_temp %>% forecast(h = 30) %>% autoplot(subset_weather)+
  labs(title = "ETS Forecast for Temperature")

```

```{r ETS_temperature_before_covid}

#Modelling ets automatically
ets_temp_covid <- weather_train_covid %>%
  model(ETS(temp))

#Getting the Accuracy scores for model
report(ets_temp_covid)

#Forecasting using the model, forecasting one month
forecast_ets_temp_covid <- ets_temp %>% forecast(h = 30)

#Graphing the forecast
ets_temp_covid %>% forecast(h = 30) %>% autoplot(subset_weather_covid)+
  labs(title = "ETS Forecast for Temperature before covid")

```

As seen above, ETS is unusable for forecasting temperature. Maybe SARIMA would be more able to capture seasonality

### Forecasting Predictors with SARIMA

```{r SARIMA_temperature_with_covid}
#Modelling ARIMA automatically
sarima_temp <- weather_train_set %>%
  model(ARIMA(temp ~ humidity, stepwise = FALSE)) 

#Getting the Accuracy scores for model
report(sarima_temp)

#Forecasting using the models
forecast_sarima_temp <- sarima_temp %>% forecast(weather_test_set)

#Graphing the forecast
autoplot(forecast_sarima_temp) +
  geom_line(aes(x=weather_test_set$date, y=weather_test_set$temp)) +
  geom_line(aes(x=subset_weather$date, y=subset_weather$temp))+
  ggtitle("Temperature forecast - SARIMAX with Humidity predictor")


forecast_table_sarimax_temp = as.data.frame(forecast_tslm_temp) %>% select(date, .mean) %>% rename(temp_SARIMAX = .mean)
head(forecast_table_sarimax_temp)


```
SARIMA does better in predicting an increase of temperatures starting in Spring
```{r SARIMA_temperature_before_covid}
# #Modelling ARIMA automatically
# sarima_temp_covid <- weather_train_covid %>%
#   model(ARIMA(temp ~ humidity, stepwise = FALSE)) 
# 
# #Getting the Accuracy scores for model
# report(sarima_temp_covid)
# 
# #Forecasting using the models
# forecast_sarima_temp_covid <- sarima_temp_covid %>% forecast(new_data = weather_test_covid)
# 
# # #Graphing the forecast
# # sarima_temp_covid %>% forecast(weather_test_covid) %>% autoplot(subset_weather_covid)+
# #   labs(title = "SARIMA Forecast for Temperature before COVID with Humidity")
# 
# autoplot(forecast_sarima_temp_covid) +
#   geom_line(aes(x=weather_test_covid$date, y=weather_test_covid$temp)) +
#   geom_line(aes(x=subset_weather_covid$date, y=subset_weather_covid$temp))+
#   ggtitle("Temperature forecast - SARIMA")

```
As we can see, SARIMA does help in predicting an increase in temperature coming close to Spring. Let's try forecasting a longer period to confirm that it captures the seasonality

### Forecasting predictors using TSLM (regression)

I will finally use a TSLM model to forecast the weather data, starting with predicting the temperature based on the rest of the weather data values.

```{r TSLM_temperature_with_covid}
#TEMPERATURE

#Fitting the TSLM model
temp_tslm <- weather_test_set %>%
  model(tslm = TSLM(temp ~ humidity))
report(temp_tslm)

#Plotting the TSLM model
augment(temp_tslm) %>%
  ggplot(aes(x = date)) +
  geom_line(aes(y = temp, colour = "Data")) +
  geom_line(aes(y = .fitted, colour = "Fitted")) + xlab("Year") + ylab(NULL) +
  guides(colour = guide_legend(title=NULL)) +
  labs(title = "TSLM to predict temperature")

#comparing the TSLM model with actual data to see how good/bad it is
augment(temp_tslm) %>%
  ggplot(aes(x = temp, y = .fitted)) + geom_point() +
  ylab("Fitted (predicted values)") + xlab("Data (actual values)") +
  geom_abline(intercept = 0, slope = 1) +
  labs(title = "Fitted values vs Actual data Temperature- Looks not amazing")

#Getting the metrics
glance(temp_tslm) %>% dplyr::select(adj_r_squared, CV, AIC, AICc, BIC)

#Getting the Forecast for March 2023:
forecast_tslm_temp <- temp_tslm %>% forecast(new_data = weather_test_set)
autoplot(forecast_tslm_temp) +
  geom_line(aes(x=weather_test_set$date, y=weather_test_set$temp)) +
  geom_line(aes(x=subset_weather$date, y=subset_weather$temp))+
  ggtitle("Temperature forecast - TSLM with Humidity predictor")


#Extracting the forecast values:
forecast_table_tslm_temp <- as.data.frame(forecast_tslm_temp)
forecast_table_tslm_temp <- forecast_table_tslm_temp %>% select(date,.mean)

#change col names
colnames(forecast_table_tslm_temp)[1] <- "date"
colnames(forecast_table_tslm_temp)[2] <- "Temp_TSLM"

head(forecast_table_tslm_temp)
```
Since this looks much better than what I was getting using SARIMA and ETS, I will use TSLM to forecast all of my predictor variables. At the moment, I don't have a metric to compare TSLM with ETS and SARIMA, I simply use my intuition based on the graph produced above. It clearly captured the usual drop in temperatures in November, whereas ETS and SARIMA produced a flat forecast.

```{r TSLM_temperature_before_covid}
# 
# #TEMPERATURE
# #Fitting the TSLM model
# temp_tslm_covid <- weather_test_covid %>%
#   model(tslm = TSLM(temp ~ humidity))
# report(temp_tslm_covid)
# 
# #Plotting the TSLM model
# augment(temp_tslm_covid) %>%
#   ggplot(aes(x = date)) +
#   geom_line(aes(y = temp, colour = "Data")) +
#   geom_line(aes(y = .fitted, colour = "Fitted")) + xlab("Year") + ylab(NULL) +
#   guides(colour = guide_legend(title=NULL)) +
#   labs(title = "TSLM to predict temperature")
# 
# #comparing the TSLM model with actual data to see how good/bad it is
# augment(temp_tslm_covid) %>%
#   ggplot(aes(x = temp, y = .fitted)) + geom_point() +
#   ylab("Fitted (predicted values)") + xlab("Data (actual values)") +
#   geom_abline(intercept = 0, slope = 1) +
#   labs(title = "Fitted values vs Actual data Temperature- Looks not amazing")
# 
# #Getting the metrics
# glance(temp_tslm_covid) %>% dplyr::select(adj_r_squared, CV, AIC, AICc, BIC)
# 
# #Getting the Forecast for November 2022:
# forecast_tslm_temp_covid <- temp_tslm_covid %>% forecast(new_data = weather_test_covid)
# #forecast_tslm_temp_covid %>% autoplot(subset_weather_covid) + labs(title = "TSLM forecast for Temperature before COVID")
# autoplot(forecast_tslm_temp_covid) +
#   geom_line(aes(x=weather_test_covid$date, y=weather_test_covid$temp)) +
#   geom_line(aes(x=subset_weather_covid$date, y=subset_weather_covid$temp))+
#   ggtitle("Temperature forecast - TSLM with Humidity predictor")
# 
# #Extracting the forecast values:
# forecast_table_tslm_temp_covid <- as.data.frame(forecast_tslm_temp_covid)
# forecast_table_tslm_temp_covid <- forecast_table_tslm_temp_covid %>% select(date,.mean)
# 
# #change col names
# colnames(forecast_table_tslm_temp_covid)[1] <- "Date"
# colnames(forecast_table_tslm_temp_covid)[2] <- "Temp_TSLM"

```

```{r tslm_windspeed_with_covid}
# 
# #WINDSPEED 
# #Fitting the TSLM model
# windspeed_tslm <- weather_train_set %>%
#   model(tslm = TSLM(windspeed ~ temp))
# report(windspeed_tslm)
# 
# #Plotting the TSLM model
# augment(windspeed_tslm) %>%
#   ggplot(aes(x = date)) +
#   geom_line(aes(y = windspeed, colour = "Data")) +
#   geom_line(aes(y = .fitted, colour = "Fitted")) + xlab("Year") + ylab(NULL) +
#   guides(colour = guide_legend(title=NULL)) +
#   labs(title = "TSLM to predict windspeed")
# 
# 
# #comparing the TSLM model with actual data to see how good/bad it is 
# augment(windspeed_tslm) %>%
#   ggplot(aes(x = windspeed, y = .fitted)) + geom_point() +
#   ylab("Fitted (predicted values)") + xlab("Data (actual values)") +
#   geom_abline(intercept = 0, slope = 1) +
#   labs(title = "Fitted values vs Actual data Windspeed")
# 
# #Getting the metrics 
# glance(windspeed_tslm) %>% dplyr::select(adj_r_squared, CV, AIC, AICc, BIC)
# 
# #Getting the Forecast for November 2022: 
# forecast_tslm_windspeed <- windspeed_tslm %>% forecast(new_data = weather_test_set, h = 30) 
# forecast_tslm_windspeed %>% autoplot(subset_weather) + labs(title = "TSLM forecast for Windspeed")
# 
# #Extracting the forecast values: 
# forecast_table_tslm_windspeed <- as.data.frame(forecast_tslm_windspeed)
# forecast_table_tslm_windspeed <- forecast_table_tslm_windspeed %>% select(date,.mean)
# 
# #change col names
# colnames(forecast_table_tslm_windspeed)[1] <- "Date"
# colnames(forecast_table_tslm_windspeed)[2] <- "Windspeed TSLM"

```
```{r tslm_windspeed_before_covid}
# 
# #WINDSPEED 
# #Fitting the TSLM model
# windspeed_tslm_covid <- weather_train_covid %>%
#   model(tslm = TSLM(windspeed ~ temp))
# report(windspeed_tslm_covid) 
# 
# #Plotting the TSLM model
# augment(windspeed_tslm_covid) %>%
#   ggplot(aes(x = date)) +
#   geom_line(aes(y = windspeed, colour = "Data")) +
#   geom_line(aes(y = .fitted, colour = "Fitted")) + xlab("Year") + ylab(NULL) +
#   guides(colour = guide_legend(title=NULL)) +
#   labs(title = "TSLM to predict windspeed")
# 
# 
# #comparing the TSLM model with actual data to see how good/bad it is 
# augment(windspeed_tslm_covid) %>%
#   ggplot(aes(x = windspeed, y = .fitted)) + geom_point() +
#   ylab("Fitted (predicted values)") + xlab("Data (actual values)") +
#   geom_abline(intercept = 0, slope = 1) +
#   labs(title = "Fitted values vs Actual data Windspeed")
# 
# #Getting the metrics 
# glance(windspeed_tslm_covid) %>% dplyr::select(adj_r_squared, CV, AIC, AICc, BIC)
# 
# #Getting the Forecast for November 2022: 
# forecast_tslm_windspeed_covid <- windspeed_tslm_covid %>% forecast(new_data = weather_test_covid, h = 30) 
# forecast_tslm_windspeed_covid %>% autoplot(subset_weather_covid) + labs(title = "TSLM forecast for Windspeed before COVID")
# 
# #Extracting the forecast values: 
# forecast_table_tslm_windspeed_covid <- as.data.frame(forecast_tslm_windspeed_covid)
# forecast_table_tslm_windspeed_covid <- forecast_table_tslm_windspeed_covid %>% select(date,.mean)
# 
# #change col names
# colnames(forecast_table_tslm_windspeed_covid)[1] <- "Date"
# colnames(forecast_table_tslm_windspeed_covid)[2] <- "Windspeed TSLM"

```

```{r sarima_windspeed_with_covid}
# arima_wind = weather_train_set %>% model(ARIMA(windspeed ~ temp))
# report(arima_wind)
# 
# forecast_arima_wind = forecast(arima_wind, new_data = weather_test_set)
# #Graphing the forecast
# forecast_arima_wind %>% autoplot(subset_weather)+
#   labs(title = "SARIMA Forecast for Windspeed")
```



At this point we have forecasts for all of our weather covariates: temperature, precipitation, windspeed and humidity. 

I decided to use TSLM because ETS and SARIMA produced complete garbage.