## Models using Weather variables

Here's what our flights-weather dataset looks like: 

```{r}
Full_Data = Flights_Holidays %>% as_tsibble(index = date)
Full_Data_Until_Covid = Full_Data[Full_Data$date <= as.Date("2020-01-31"),] %>% as_tsibble(index= date)

datatable(Full_Data, options = list(pageLength = 6))

weather_train_set = Full_Data[Full_Data$date < as.Date("2023-01-01"),]
weather_test_set = Full_Data[Full_Data$date >= as.Date("2023-01-01"),]
weather_train_covid = Full_Data_Until_Covid[Full_Data_Until_Covid$date < as.Date("2020-01-01"),]
weather_test_covid = Full_Data_Until_Covid[Full_Data_Until_Covid$date >= as.Date("2020-01-01"),]
subset_weather = weather_train_set[weather_train_set$date >= as.Date("2022-11-01"),]
subset_weather_covid = weather_train_covid[weather_train_covid$date >= as.Date("2019-11-01"),]

```

### Forecasting Predictors with ETS 

```{r ETS_temperature_with_covid}

#Modelling ets automatically
ets_temp <- weather_train_set %>%
  model(ETS(temp))

#Getting the Accuracy scores for model
report(ets_temp)

#Forecasting using the model, forcating one month
forecast_ets_temp <- ets_temp %>% forecast(h = 30)

#Graphing the forecast
ets_temp %>% forecast(h = 30) %>% autoplot(subset_weather)+
  labs(title = "ETS Forecast for Temperature")

```
```{r ETS_temperature_before_covid}

#Modelling ets automatically
ets_temp_covid <- weather_train_covid %>%
  model(ETS(temp))

#Getting the Accuracy scores for model
report(ets_temp_covid)

#Forecasting using the model, forecasting one month
forecast_ets_temp_covid <- ets_temp %>% forecast(h = 30)

#Graphing the forecast
ets_temp_covid %>% forecast(h = 30) %>% autoplot(subset_weather_covid)+
  labs(title = "ETS Forecast for Temperature before covid")

```


Here we can see that the ETS model is complete garbage at predicting that temperatures will drop in November.

### Forecasting Predictors with SARIMA

```{r SARIMA_temperature_with_covid}
#Modelling ARIMA automatically
arima_temp <- weather_train_set %>%
  select(temp) %>%
  model(ARIMA(temp, stepwise = FALSE)) 

#Getting the Accuracy scores for model
report(arima_temp) #ARIMA(3,1,1)(2,0,0)[7]

#Forecasting using the models
forecast_arima_temp <- arima_temp %>% forecast(h = 30)

#Graphing the forecast
arima_temp %>% forecast(h = 30) %>% autoplot(subset_weather)+
  labs(title = "SARIMA Forecast for Temperature")

```
```{r SARIMA_temperature_before_covid}
#Modelling ARIMA automatically
arima_temp_covid <- weather_train_covid %>%
  select(temp) %>%
  model(ARIMA(temp, stepwise = FALSE)) 

#Getting the Accuracy scores for model
report(arima_temp_covid) #ARIMA(3,1,1)(2,0,0)[7]

#Forecasting using the models
forecast_arima_temp <- arima_temp_covid %>% forecast(h = 30)

#Graphing the forecast
arima_temp_covid %>% forecast(h = 30) %>% autoplot(subset_weather_covid)+
  labs(title = "SARIMA Forecast for Temperature before COVID")

```

Here we can see that the SARIMA model is complete garbage at predicting that temperatures will drop in November.

### Forecasting predictors using TSLM (regression)

I will finally use a TSLM model to forecast the weather data, starting with predicting the temperature based on the rest of the weather data values.

```{r TSLM_temperature_with_covid}
#TEMPERATURE 

#Fitting the TSLM model
temp_tslm <- weather_test_set %>%
  model(tslm = TSLM(temp ~ precip + windspeed + humidity))
report(temp_tslm)

#Plotting the TSLM model
augment(temp_tslm) %>%
  ggplot(aes(x = date)) +
  geom_line(aes(y = temp, colour = "Data")) +
  geom_line(aes(y = .fitted, colour = "Fitted")) + xlab("Year") + ylab(NULL) +
  guides(colour = guide_legend(title=NULL)) +
  labs(title = "TSLM to predict temperature")

#comparing the TSLM model with actual data to see how good/bad it is 
augment(temp_tslm) %>%
  ggplot(aes(x = temp, y = .fitted)) + geom_point() +
  ylab("Fitted (predicted values)") + xlab("Data (actual values)") +
  geom_abline(intercept = 0, slope = 1) +
  labs(title = "Fitted values vs Actual data Temperature- Looks not amazing")

#Getting the metrics 
glance(temp_tslm) %>% dplyr::select(adj_r_squared, CV, AIC, AICc, BIC)

#Getting the Forecast for November 2022: 
forecast_tslm_temp <- temp_tslm %>% forecast(new_data = weather_test_set, h = 30) 
forecast_tslm_temp %>% autoplot(subset_weather) + labs(title = "TSLM forecast for Temperature")

#Extracting the forecast values: 
forecast_table_tslm_temp <- as.data.frame(forecast_tslm_temp)
forecast_table_tslm_temp <- forecast_table_tslm_temp %>% select(date,.mean)

#change col names
colnames(forecast_table_tslm_temp)[1] <- "Date"
colnames(forecast_table_tslm_temp)[2] <- "Temp TSLM"
```
Since this looks much better than what I was getting using SARIMA and ETS, I will use TSLM to forecast all of my predictor variables. At the moment, I don't have a metric to compare TSLM with ETS and SARIMA, I simply use my intuition based on the graph produced above. It clearly captured the usual drop in temperatures in November, whereas ETS and SARIMA produced a flat forecast.

```{r TSLM_temperature_before_covid}
#TEMPERATURE 

#Fitting the TSLM model
temp_tslm_covid <- weather_test_covid %>%
  model(tslm = TSLM(temp ~ precip + windspeed + humidity))
report(temp_tslm_covid)

#Plotting the TSLM model
augment(temp_tslm_covid) %>%
  ggplot(aes(x = date)) +
  geom_line(aes(y = temp, colour = "Data")) +
  geom_line(aes(y = .fitted, colour = "Fitted")) + xlab("Year") + ylab(NULL) +
  guides(colour = guide_legend(title=NULL)) +
  labs(title = "TSLM to predict temperature")

#comparing the TSLM model with actual data to see how good/bad it is 
augment(temp_tslm_covid) %>%
  ggplot(aes(x = temp, y = .fitted)) + geom_point() +
  ylab("Fitted (predicted values)") + xlab("Data (actual values)") +
  geom_abline(intercept = 0, slope = 1) +
  labs(title = "Fitted values vs Actual data Temperature- Looks not amazing")

#Getting the metrics 
glance(temp_tslm_covid) %>% dplyr::select(adj_r_squared, CV, AIC, AICc, BIC)

#Getting the Forecast for November 2022: 
forecast_tslm_temp_covid <- temp_tslm_covid %>% forecast(new_data = weather_test_covid, h = 30) 
forecast_tslm_temp_covid %>% autoplot(subset_weather_covid) + labs(title = "TSLM forecast for Temperature before COVID")

#Extracting the forecast values: 
forecast_table_tslm_temp_covid <- as.data.frame(forecast_tslm_temp_covid)
forecast_table_tslm_temp_covid <- forecast_table_tslm_temp_covid %>% select(date,.mean)

#change col names
colnames(forecast_table_tslm_temp_covid)[1] <- "Date"
colnames(forecast_table_tslm_temp_covid)[2] <- "Temp TSLM"
```

```{r}
# 
# #PRECIPITATION 
# #Fitting the TSLM model
# precip_tslm <- ts_data %>%
#   model(tslm = TSLM(precip ~ temp + windspeed + humidity))
# report(precip_tslm)
# 
# #Plotting the TSLM model
# augment(precip_tslm) %>%
#   ggplot(aes(x = Date)) +
#   geom_line(aes(y = precip, colour = "Data")) +
#   geom_line(aes(y = .fitted, colour = "Fitted")) + xlab("Year") + ylab(NULL) +
#   guides(colour = guide_legend(title=NULL)) +
#   labs(title = "TSLM to predict precipitation")
# 
# 
# #comparing the TSLM model with actual data to see how good/bad it is 
# augment(precip_tslm) %>%
#   ggplot(aes(x = precip, y = .fitted)) + geom_point() +
#   ylab("Fitted (predicted values)") + xlab("Data (actual values)") +
#   geom_abline(intercept = 0, slope = 1) +
#   labs(title = "Fitted values vs Actual data Precipitation")
# 
# #Getting the metrics 
# glance(precip_tslm) %>% dplyr::select(adj_r_squared, CV, AIC, AICc, BIC)
# 
# #Getting the Forecast for November 2022: 
# forecast_tslm_precip <- precip_tslm %>% forecast(new_data = ts_data_weather_te, h = 15) 
# forecast_tslm_precip %>% autoplot(ts_data_weather_last6) + labs(title = "TSLM forecast for Precipitation")
# 
# #Extracting the forecast values: 
# forecast_table_tslm_precip <- as.data.frame(forecast_tslm_precip)
# forecast_table_tslm_precip <- forecast_table_tslm_precip %>% select(Date,.mean)
# 
# #change col names
# colnames(forecast_table_tslm_precip)[1] <- "Date"
# colnames(forecast_table_tslm_precip)[2] <- "Precip TSLM"

```

```{r tslm_windspeed_with_covid}


#WINDSPEED 
#Fitting the TSLM model
windspeed_tslm <- weather_train_set %>%
  model(tslm = TSLM(windspeed ~ temp + precip + humidity))
report(windspeed_tslm)

#Plotting the TSLM model
augment(windspeed_tslm) %>%
  ggplot(aes(x = date)) +
  geom_line(aes(y = windspeed, colour = "Data")) +
  geom_line(aes(y = .fitted, colour = "Fitted")) + xlab("Year") + ylab(NULL) +
  guides(colour = guide_legend(title=NULL)) +
  labs(title = "TSLM to predict windspeed")


#comparing the TSLM model with actual data to see how good/bad it is 
augment(windspeed_tslm) %>%
  ggplot(aes(x = windspeed, y = .fitted)) + geom_point() +
  ylab("Fitted (predicted values)") + xlab("Data (actual values)") +
  geom_abline(intercept = 0, slope = 1) +
  labs(title = "Fitted values vs Actual data Windspeed")

#Getting the metrics 
glance(windspeed_tslm) %>% dplyr::select(adj_r_squared, CV, AIC, AICc, BIC)

#Getting the Forecast for November 2022: 
forecast_tslm_windspeed <- windspeed_tslm %>% forecast(new_data = weather_test_set, h = 30) 
forecast_tslm_windspeed %>% autoplot(subset_weather) + labs(title = "TSLM forecast for Windspeed")

#Extracting the forecast values: 
forecast_table_tslm_windspeed <- as.data.frame(forecast_tslm_windspeed)
forecast_table_tslm_windspeed <- forecast_table_tslm_windspeed %>% select(date,.mean)

#change col names
colnames(forecast_table_tslm_windspeed)[1] <- "Date"
colnames(forecast_table_tslm_windspeed)[2] <- "Windspeed TSLM"

```
```{r tslm_windspeed_before_covid}


#WINDSPEED 
#Fitting the TSLM model
windspeed_tslm_covid <- weather_train_covid %>%
  model(tslm = TSLM(windspeed ~ temp + precip + humidity))
report(windspeed_tslm_covid)

#Plotting the TSLM model
augment(windspeed_tslm_covid) %>%
  ggplot(aes(x = date)) +
  geom_line(aes(y = windspeed, colour = "Data")) +
  geom_line(aes(y = .fitted, colour = "Fitted")) + xlab("Year") + ylab(NULL) +
  guides(colour = guide_legend(title=NULL)) +
  labs(title = "TSLM to predict windspeed")


#comparing the TSLM model with actual data to see how good/bad it is 
augment(windspeed_tslm_covid) %>%
  ggplot(aes(x = windspeed, y = .fitted)) + geom_point() +
  ylab("Fitted (predicted values)") + xlab("Data (actual values)") +
  geom_abline(intercept = 0, slope = 1) +
  labs(title = "Fitted values vs Actual data Windspeed")

#Getting the metrics 
glance(windspeed_tslm_covid) %>% dplyr::select(adj_r_squared, CV, AIC, AICc, BIC)

#Getting the Forecast for November 2022: 
forecast_tslm_windspeed_covid <- windspeed_tslm_covid %>% forecast(new_data = weather_test_covid, h = 30) 
forecast_tslm_windspeed_covid %>% autoplot(subset_weather_covid) + labs(title = "TSLM forecast for Windspeed before COVID")

#Extracting the forecast values: 
forecast_table_tslm_windspeed_covid <- as.data.frame(forecast_tslm_windspeed_covid)
forecast_table_tslm_windspeed_covid <- forecast_table_tslm_windspeed_covid %>% select(date,.mean)

#change col names
colnames(forecast_table_tslm_windspeed_covid)[1] <- "Date"
colnames(forecast_table_tslm_windspeed_covid)[2] <- "Windspeed TSLM"

```

```{r}
# 
# 
# #HUMIDITY 
# #Fitting the TSLM model
# humidity_tslm <- ts_data %>%
#   model(tslm = TSLM(humidity ~ temp + precip + windspeed))
# report(humidity_tslm)
# 
# #Plotting the TSLM model
# augment(humidity_tslm) %>%
#   ggplot(aes(x = Date)) +
#   geom_line(aes(y = humidity, colour = "Data")) +
#   geom_line(aes(y = .fitted, colour = "Fitted")) + xlab("Year") + ylab(NULL) +
#   guides(colour = guide_legend(title=NULL)) +
#   labs(title = "TSLM to predict humidity")
# 
# 
# #comparing the TSLM model with actual data to see how good/bad it is 
# augment(humidity_tslm) %>%
#   ggplot(aes(x = humidity, y = .fitted)) + geom_point() +
#   ylab("Fitted (predicted values)") + xlab("Data (actual values)") +
#   geom_abline(intercept = 0, slope = 1) +
#   labs(title = "Fitted values vs Actual data Humidity")
# 
# #Getting the metrics 
# glance(humidity_tslm) %>% dplyr::select(adj_r_squared, CV, AIC, AICc, BIC)
# 
# #Getting the Forecast for November 2022: 
# forecast_tslm_humidity <- humidity_tslm %>% forecast(new_data = ts_data_weather_te, h = 15) 
# forecast_tslm_humidity %>% autoplot(ts_data_weather_last6) + labs(title = "TSLM forecast for Humidity")
# 
# #Extracting the forecast values: 
# forecast_table_tslm_humidity <- as.data.frame(forecast_tslm_humidity)
# forecast_table_tslm_humidity <- forecast_table_tslm_humidity %>% select(Date,.mean)
# 
# #change col names
# colnames(forecast_table_tslm_humidity)[1] <- "Date"
# colnames(forecast_table_tslm_humidity)[2] <- "Humidity TSLM"

```

At this point we have forecasts for all of our weather covariates: temperature, precipitation, windspeed and humidity. 

I decided to use TSLM because ETS and SARIMA produced complete garbage.