
# Models 

We decided to split the modeling section into two parts. In the first part, we will try to forecast the number of daily flights, for the following 12 months, only based on the price of oil. In the second part, we will try to forecast the number of daily flights using the weather variables, for the next two weeks. We therefore decided not to mix these oil price and weather variables together. 

We made this decision, because as we saw in our EDA the price of oil is not as volatile as our weather variables like the temperature. More specifically, we don't expect the price of oil to change in a matter of two weeks. At the same time, weather data is extremely hard to forecast, and it doesn't really make sense to forecast weather data for a year from now. Because of these two reasons, we decided not to mix the two, and instead split the modeling part of our project into two parts.

## Models using the Oil price variable

Here's what our flights-oil dataset looks like: 

```{r}
datatable(ts_data_oil, options = list(pageLength = 6))

```

### Let's start by forecasting the Price variable

Before using it, we need to forecast the price variable from June 2022 until November 2022. To do so, we will use different models (ETS, SARIMA, Regression), and then we will compare which one was more accurate.

### ETS Forecast 

For training the model, I will use the ts_data_oil_tr dataset.

```{r}

#Modelling ets automatically
ets_oil <- ts_data_oil_tr %>%
  model(ETS(Price))

#Getting the Accuracy scores for model
report(ets_oil)

#Forecasting using the model -> 317 days between 1st of jan and 13th of November
forecast_ets_oil <- ets_oil %>% forecast(h = 317)

#Graphing the forecast
ets_oil %>% forecast(h = 317) %>% autoplot(ts_data_oil)+
  labs(title = "ETS Forecast for Oil prices")

#Extracting the forecasted values
forecast_table_oil <- data.frame(Model = c("ETS"),
                              Forecast_Values = c(forecast_ets_oil))

forecast_table_oil <- forecast_table_oil %>% select(Forecast_Values.Date,Forecast_Values..mean)

colnames(forecast_table_oil)[1] <- "Date"
colnames(forecast_table_oil)[2] <- "OIL ETS"

```

The ETS looks terrible, there must be something wrong. 

### SARIMA Forecast

```{r}
#Modelling ARIMA automatically
arima_oil <- ts_data_oil_tr %>%
  select(Price) %>%
  model(ARIMA(Price, stepwise = FALSE)) 

#Getting the Accuracy scores for model
report(arima_oil) #ARIMA(3,1,1)(0,0,2)[7]

#Forecasting using the models
forecast_arima_oil <- arima_oil %>% forecast(h = 317)

#Graphing the forecast
arima_oil %>% forecast(h = 317) %>% autoplot(ts_data_oil)+
  labs(title = "SARIMA Forecast for Oil Prices")

#Extracting the forecasted values
forecast_table_oil_arima <- data.frame(Model = c("SARIMA"),
                              Forecast_Values = c(forecast_arima_oil))

forecast_table_oil_arima <- forecast_table_oil_arima %>% select(Forecast_Values.Date,Forecast_Values..mean)

colnames(forecast_table_oil_arima)[1] <- "Date"
colnames(forecast_table_oil_arima)[2] <- "OIL SARIMA"
```

How can this be so bad? 

The explanation is that this data is very unpredictable and difficult to forecast. Also, anyone looking at this data before 2022 would have absolutely no idea that the oil prices would explode like this, because of the war... That was an extreme, unpredictable event that caused oil prices to skyrocket... I thought it would be good to include the this oil price variable, because it's very correlated with the number of flights, but it seems like it's not possible to predict it accurately... 

### Regression Forecast 











