
# Forecasting Flights

After having forecasted and selected our covariates, it's time to forecast the number of flights. 

First, let's create a table that contains all the forecasts from our covariates: 

```{r}
#Importing actual data
Full_Data = Flights_Holidays

#Creating datasets with predicted covariates
Full_Data_With_Forecasts_tslm = merge(Full_Data, forecast_table_tslm_temp, by = "date", all.x = TRUE)
Full_Data_With_Forecasts_sarimax = merge(Full_Data, forecast_table_sarimax_temp, by = "date", all.x = TRUE)
Full_Data_With_Forecasts_tslm$temp = ifelse(is.na(Full_Data_With_Forecasts_tslm$Temp_TSLM),
                                            Full_Data_With_Forecasts_tslm$temp, 
                                            Full_Data_With_Forecasts_tslm$Temp_TSLM)

Full_Data_With_Forecasts_sarimax$temp = ifelse(is.na(Full_Data_With_Forecasts_sarimax$temp_SARIMAX),
                                            Full_Data_With_Forecasts_sarimax$temp, 
                                            Full_Data_With_Forecasts_sarimax$temp_SARIMAX)

#Filling dataset from TSLM predictions
rows_to_fill = Full_Data_With_Forecasts_tslm$date >= as.Date("2023-03-01")
Full_Data_With_Forecasts_tslm$departures[rows_to_fill] = NA

#We want to predict the month of march
forecast_train_set_tslm = Full_Data_With_Forecasts_tslm %>% filter(date < as.Date("2023-03-01")) %>% as_tsibble(index = date)
forecast_test_set_tslm = Full_Data_With_Forecasts_tslm %>% filter(date >= as.Date("2023-03-01")) %>% as_tsibble(index = date)

# #Filling dataset with Sarimax predictions
# rows_to_fill = Full_Data_With_Forecasts_sarimax$date >= as.Date("2023-03-01")
# Full_Data_With_Forecasts_sarimax$departures[rows_to_fill] = NA
# #We want to predict the month of march
# forecast_train_set_arima = Full_Data_With_Forecasts_sarimax %>% filter(date < as.Date("2023-03-01")) %>% as_tsibble(index = date)
# forecast_test_set_arima = Full_Data_With_Forecasts_sarimax %>% filter(date >= as.Date("2023-03-01")) %>% as_tsibble(index = date)

#Reference data with actual values
holidays_total = Full_Data %>% mutate(is_holiday = as.factor(is_holiday)) %>% as_tsibble(index = date)
holidays_train_set = holidays_total %>% filter(date < as.Date("2023-03-01"))
holidays_test_set = holidays_total %>% filter(date >= as.Date("2023-03-01"))

```

```{r before_covid}
Full_Data_Until_Covid = Full_Data[Full_Data$date <= as.Date("2020-04-30"),]
#Creating datasets with predicted covariates
Data_to_Covid_With_Forecasts_tslm = merge(Full_Data_Until_Covid, forecast_table_tslm_temp_covid, by = "date", all.x = TRUE)
Data_to_Covid_With_Forecasts_sarimax = merge(Full_Data_Until_Covid, forecast_table_sarimax_temp_covid, by = "date", all.x = TRUE)
Data_to_Covid_With_Forecasts_tslm$temp = ifelse(is.na(Data_to_Covid_With_Forecasts_tslm$temp_TSLM),
                                            Data_to_Covid_With_Forecasts_tslm$temp, 
                                            Data_to_Covid_With_Forecasts_tslm$temp_TSLM)

Data_to_Covid_With_Forecasts_sarimax$temp = ifelse(is.na(Data_to_Covid_With_Forecasts_sarimax$temp_SARIMAX),
                                            Data_to_Covid_With_Forecasts_sarimax$temp, 
                                            Data_to_Covid_With_Forecasts_sarimax$temp_SARIMAX)

#Data before covid
forecast_train_set_tslm_covid = Data_to_Covid_With_Forecasts_tslm %>% filter(date < as.Date("2020-03-15")) %>% as_tsibble(index = date)
forecast_test_set_tslm_covid = Data_to_Covid_With_Forecasts_tslm %>% filter(date >= as.Date("2020-03-15")) %>% as_tsibble(index = date)


holidays_total_covid = Full_Data %>% mutate(is_holiday = as.factor(is_holiday)) %>% as_tsibble(index = date)
holidays_train_covid = holidays_total_covid %>% filter(date < as.Date("2020-03-15"))
holidays_test_covid = holidays_total_covid %>% filter(date >= as.Date("2020-03-15"))

```


```{r Sarimax_with_forecastedV_with_covid}
#With COVID and forecasted values of holidays and temp
ARIMA_with_forecasted_H_T = forecast_train_set_tslm %>%
  model(ARIMA(departures ~ is_holiday + temp))

#Accuracy score with forecasted values
report(ARIMA_with_forecasted_H_T)

#Forecasting values of temperature for month of march
ARIMA_forecasted_D_H_T = ARIMA_with_forecasted_H_T %>% forecast(new_data = forecast_test_set_tslm)

#Plotting results with forecasted temperatures and holidays
subset_data_with_covid = forecast_train_set_tslm[forecast_train_set_tslm$date >= as.Date("2022-12-01"),]
autoplot(ARIMA_forecasted_D_H_T) +
  geom_line(aes(x=forecast_test_set_tslm$date, y=forecast_test_set_tslm$departures)) +
  geom_line(aes(x=subset_data_with_covid$date, y=subset_data_with_covid$departures)) +
  geom_line(aes(x=holidays_test_set$date, y=holidays_test_set$departures)) +
  ggtitle("ARIMA model for whole time frame with forecasted T + H")

```
```{r Sarimax_with_forecastedV_before_covid}
#With COVID and forecasted values of holidays and temp
ARIMA_with_forecasted_H_T_covid = forecast_train_set_tslm_covid %>%
  model(ARIMA(departures ~ is_holiday + temp))

#Accuracy score with forecasted values
report(ARIMA_with_forecasted_H_T_covid)

#Forecasting values of temperature for month of march
ARIMA_forecasted_D_H_T_covid = ARIMA_with_forecasted_H_T_covid %>% forecast(new_data = forecast_test_set_tslm_covid)

#Plotting results with forecasted temperatures and holidays
subset_data_covid = forecast_train_set_tslm_covid[forecast_train_set_tslm_covid$date >= as.Date("2019-10-01") & forecast_train_set_tslm_covid$date < as.Date("2020-05-01"),]
autoplot(ARIMA_forecasted_D_H_T_covid) +
  geom_line(aes(x=forecast_test_set_tslm_covid$date, y=forecast_test_set_tslm_covid$departures)) +
  geom_line(aes(x=subset_data_covid$date, y=subset_data_covid$departures)) +
  ggtitle("Departures if COVID didn't happen - ARIMA holidays + temperature")

```


```{r TSLM_with_forecastedV_with_covid}

#With covid and forecasted values of temperature and holidays using ARIMA

#With holidays and temperature
TSLM_with_forecasted_H_T = forecast_train_set_tslm %>%
  model(TSLM(departures ~ is_holiday + temp))

#Accuracy score
report(TSLM_with_forecasted_H_T)
glance(TSLM_with_forecasted_H_T) %>% dplyr::select(adj_r_squared, CV, AIC, AICc, BIC)

#Forecasting values
forecast_D_H_T = TSLM_with_forecasted_H_T %>% forecast(new_data = forecast_test_set_tslm)

#Plotting results
subset_data_with_covid = forecast_train_set_tslm[forecast_train_set_tslm$date >= as.Date("2022-12-01"),]
autoplot(forecast_D_H_T) +
  geom_line(aes(x=forecast_test_set_tslm$date, y=forecast_test_set_tslm$departures)) +
  geom_line(aes(x=subset_data_with_covid$date, y=subset_data_with_covid$departures)) +
  geom_line(aes(x=holidays_test_set$date, y=holidays_test_set$departures)) +
  ggtitle("ARIMA model for whole time frame")


```
TSLM is not useful when forecasting flights only with theses parameters 



```{r arima_model_with_covid }
# #With covid and forecasted values using TSLM
# holidays_total = Full_Data %>% mutate(is_holiday = as.factor(is_holiday)) %>% as_tsibble(index = date)
# holidays_train_set = holidays_total %>% filter(date < as.Date("2023-01-01"))
# holidays_test_set = holidays_total %>% filter(date >= as.Date("2023-01-01"))
# 
# #Without holidays
# arima_flights = holidays_train_set %>% 
#   model(ARIMA(departures, stepwise= FALSE)) %>% 
#   forecast(h = nrow(holidays_test_set))
# 
# #With holidays
# sarimax_flights_H_W_T = holidays_train_set %>% 
#   model(ARIMA(departures ~ is_holiday, stepwise= FALSE)) %>% 
#   forecast(new_data = as_tsibble(holidays_test_set))
# 
# 
# #With holidays
# sarimax_flights_H_W_T = holidays_train_set %>% 
#   model(ARIMA(departures ~ is_holiday + windspeed, stepwise= FALSE)) %>% 
#   forecast(new_data = as_tsibble(holidays_test_set))
# 
# #With holidays
# sarimax_flights_H_W_T = holidays_train_set %>% 
#   model(ARIMA(departures ~ is_holiday + temp, stepwise= FALSE)) %>% 
#   forecast(new_data = as_tsibble(holidays_test_set))
# 
# #Plotting
# subset_data_with_covid = holidays_train_set[holidays_train_set$date >= as.Date("2020-12-01"),]
# autoplot(arima_flights) +
#   geom_line(aes(x=holidays_test_set$date, y=holidays_test_set$departures)) +
#   geom_line(aes(x=subset_data_with_covid$date, y=subset_data_with_covid$departures))+
#   ggtitle("ARIMA model for whole time frame")
# 
# autoplot(sarimax_flights_H_W_T) +
#   geom_line(aes(x=holidays_test_set$date, y=holidays_test_set$departures)) +
#   geom_line(aes(x=subset_data_with_covid$date, y=subset_data_with_covid$departures))+
#   ggtitle("SARIMAX model with for holiday regressor")

```
From the graphs above we can see that the forecast is not very accurate because it hardly captures the seasonal component of our time series. This is certainly due to the fact that flights dropped severely because of covid, add a lot of noise in the model. Let's see how it performs when we try forecasting values before covid as if it didn't happen.



```{r arima_model_before_covid}
# #Before covid
# 
# # #Creating train and test sets
# # holidays_total_covid = Full_Data_Until_Covid %>% mutate(is_holiday = as.factor(is_holiday)) %>% as_tsibble(index = date)
# # holidays_train_covid = holidays_total_covid %>% filter( date < as.Date("2020-01-01"))
# # holidays_test_covid = holidays_total_covid %>% filter(date >= as.Date("2020-01-01"))
# 
# #Without holidays
# arima_flights = holidays_train_covid %>% 
#   model(ARIMA(departures, stepwise= FALSE)) %>% 
#   forecast(h = nrow(holidays_test_covid))
# 
# #With holidays
# sarimax_flights_covid_H =  %>% 
#   model(ARIMA(departures ~ is_holiday, stepwise= FALSE)) %>% 
#   forecast(new_data = as_tsibble(holidays_test_covid)) 
# 
# #With forecasted temperature
# sarimax_flights_covid_H_T = 
# 
# #Plotting
# 
# subset_data = holidays_train_covid[holidays_train_covid$date >= as.Date("2018-06-01"),]
# 
# #Without regressors
# autoplot(arima_flights) +
#   geom_line(aes(x=holidays_test_covid$date, y=holidays_test_covid$departures)) +
#   geom_line(aes(x=subset_data$date, y=subset_data$departures))+
#   ggtitle("ARIMA forecast with data until 01.01.2020")
# 
# #With regressors
# autoplot(sarimax_flights_covid_H_W_T) + 
#   geom_line(aes(x = holidays_test_covid$date, y = holidays_test_covid$departures)) +
#   geom_line(aes(x=subset_data$date, y=subset_data$departures))+
#   ggtitle("SARIMAX H-W-T with data until 01.01.2020 ") 


```
